Discussion

### 2020-03-13 doc/Discussion.rmd
notes from Wu thesis

## Typical landscape
<review the typical landscape, one of the major broad scale patterns,
exceptions
consequences, (differences in variation, Sardell paper) >

@@

## Chromatin Structure

< review loop-axis relationship,  review how interference is transmitter throught the axis  -- transformation of genetic area -- results in longer loops having stronger interference,

OUR pattern is that

>

@@

@@

1.Gamete selection
The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.
2.SACE
Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.
3.Two-locus (protect for drivers)
Asymmetric division of egg opens them up to meiotic drivers. Recombination modifiers for sex specific female landscape will decrease the chances of driving centromeres to segregate to the egg by increasing the number of crossovers and placing them across a larger area of the genome (uniform placement).
4.Difference in acentromeric spindle and SAC
A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle). (Our hypothesis is that the tension in a centrosome spindle is stronger (or more uniform) since all MT-KT are anchored at single points across cells (the two centrosomes). This will also make prediction for the spindle assembly checkpoint (SAC), which is stronger/stricter in smaller gametes with centrosomes.  (the SAC is more sensitive to achiasmate bivalents on the spindle / better at detecting a lack of tension).
Whereas in the acentrosome spindle is more diffuse (across a larger area) and the tension across bivalents are anchored by multiple MTOCs, resulting in weaker tension force or less uniform strengths across all the bivalents within the cells. The SAC in eggs seems to be weaker/leakier, that is more achiasmate bivalents are required to trigger the SAC (stop the division).

@@

>>>>

## male telomeric, female centromeric Typical landscape
<review the typical landscape, one of the major broad scale patterns,
exceptions
consequences, (differences in variation, Sardell paper) >

@@

The Sardel guppies paper was able to deduce (dientangle) that this typical landscape pattern was due to // disentangle the specific components of centromere supression and telomere bias between the male and feamel landscapes. we are not able to do that in these cases (because of inability to estimate the chromosome identity across our sample of single bivalents. )

## Chromatin Structure and interference strength

@@

Petkov paper -- states that the sexual dimorphism for interference strength is due to the chromatin structure (meiotic chromosome organization) with the increased interference strength in males being due to longer chromatin loop lengths.

@@

## 1.Gamete selection
The process where gametes are competing - so there is stronger selection for 1 form of gamete vs the other (stronger directional selection) (the gamete with more competition / the sex with more variance in reproductive fitness will have lower overall recombination rate.

<go through the squares>

this is a modifer model it abstracts the genome area.

This model predicts the evolution of heterchiasmy -- through models (that build on selection to reduce recombiantion for the sex with geater varatince in fitness effect)

this model doesn't predict that the heterochismy direction would reverse.

##2.SACE
Extension of reduction principle. Results in large blocks of genetic areas being kept together in male meiosis.

##3.Two-locus (protect for drivers)
Asymmetric division of egg opens them up to meiotic drivers. Recombination modifiers for sex specific female landscape will decrease the chances of driving centromeres to segregate to the egg by increasing the number of crossovers and placing them across a larger area of the genome (uniform placement).

## 4. Sindle   // Difference in acentromeric spindle and SAC
A fundamental difference in meiosis (MI) for males and females is the presence of the centrosome which serves as a nucleation site for MT at each pole (and changes the shape of the spindle). (Our hypothesis is that the tension in a centrosome spindle is stronger (or more uniform) since all MT-KT are anchored at single points across cells (the two centrosomes). This will also make prediction for the spindle assembly checkpoint (SAC), which is stronger/stricter in smaller gametes with centrosomes.  (the SAC is more sensitive to achiasmate bivalents on the spindle / better at detecting a lack of tension).
Whereas in the acentrosome spindle is more diffuse (across a larger area) and the tension across bivalents are anchored by multiple MTOCs, resulting in weaker tension force or less uniform strengths across all the bivalents within the cells. The SAC in eggs seems to be weaker/leakier, that is more achiasmate bivalents are required to trigger the SAC (stop the division).

<schuh ellenberg 2007 spindle

_review the correct SAC_

bivalent orientation early metaphase -- search and capture model -- when there are mulitple KT-MT attachmments, (lateral (on the side or not on centromere/KT proper), merotelic MT attachemtn from ooposite poles) and amplitelic (MT from same pole)
then biorientation process starts, --search and capature process (only amplictic Mt-KT attachments get stablize)

Aurora -- detects bad connections,

_evidence that SAC in oogenesis is leaky_

(Lane et al 2012)

whenn going through the attachment correection, bivalents are thought to be rotated via pulling from MT

Akera 2017 -- chromosomes at different ends of spindle have different densites (in reference to the cortical or central side of spindle (PB or egg))

Wu thesis -- chrm 17 with asymetric centromere str, cenromeres with large major stat drove to (polar body (opposite of drive) (lack of re-orientation, once it ended up on that side))

the biased rotation happened before spindle migratgation (from center of egg to near cortex)

when a aurora inhibitor applied -- the biased rotation of chrm 17 went away and anaphase moved up (by passed SAC as in Lane 2010)

(k-fibre the bundle MTs connected to KT Iwata-otsubo 2017) -- larger KT would provide more k0fiber binding sites

the spindle migration (from center of egg to cortext depends on actin (Li albertini 2013))
cytoplasmic streaming when spindle moves close to the cortex (yi et al 2013)

the KT and centromere seperate when MT's start pulling (trains start moving) (they does seperate whn MT inhibitors are applied)

Coretical KTs undergo greater tension than centrak KTs.

@@

aMTOC (Cliff and Schuh 2015, Schuh and ellenburg)
 since there are multiple aMTOC (not 2 centrioles) -- the MTOCs across the sides of the cells could be unven.

after NEBD the aMTOCs are fragmented and coalest into two main clusters (Clift and Schuh 2015 and Luksza 2013). 
(Kolano et al 2012 maintence of spindle)
aMTOC migration to pools is dynamic and self organized.

Wu thesis compared th density of spindles on the central and cortical halves
data suggests that more MTOC are recruited to the cortical side
